Challenges in Developing Great MCQMC Software

### 2022-07-17 MCQMC_2022_ChallengesDevelopSoftware.tex
Office Computer Updates

\usetheme{FJHSlimNoFoot169}
\setlength{\parskip}{2ex}
\setlength{\arraycolsep}{0.5ex}
\setbeamersize{text margin top=10mm}

@@

\begin{frame}{My Definition of Great MCQMC Software}

@@

\begin{frame}[allowframebreaks]{Selective History of Publicly Available MC\textbf{QMC} Software}
\vspace{-2ex}
	\begin{itemize}
		\item \textcite{LEc2017a} provides a history of \alert{random number generation}
		\item Notorious \alert{RANDU} from the 1960s \parencite{RANDU}, which fails the spectral test
		\item \alert{VEGAS} \parencite{Lep78a, Lep21a} importance sampling Monte Carlo algorithm favored by physicists
		\item \alert{ACM} low discrepancy point generators \parencite{BraFox88,BraFoxNie92,HonHic00a}
		\item \alert{FinDer} \parencite{PasTra95,FinDer} and \alert{BRODA} \parencite{BRODA20a} targeting quantitative finance
		\item Korobov cubature and scrambled Sobol' generators in \alert{NAG} \parencite{NAG27} for decades		
		\item Scrambled Sobol' and Halton generators in \alert{MATLAB} \parencite{MAT9.13} since 2008, and fixed a few years later
		\item \alert{BUGS} \parencite{BUGSBook, BUGSweb} and \alert{Stan} \parencite{STAN} for Markov Chain Monte Carlo (MCMC)
		\item \alert{LatMRG} \parencite{LEcCou97}, \alert{RNGStreams} \parencite{LEcEtal02},  \alert{SSJ} \parencite{LEc2002a,SSJ}, and \alert{LatNetBuilder} \parencite{LatNet}
		\item Low discrepancy generators \parencite{FriKel02,FriKelweb}, \alert{SamplePack} \parencite{SamplePack}, \textcite{GruWeb},  and \alert{MatBuilder} \parencite{paulin2022}
		\item \alert{Fast CBC}, \alert{Magic Point Shop}, \alert{QMC4PDE} and other code since 2004  at \textcite{NuyWeb}
		\item Multi-level software \parencite{GilesSoft,GilesQSoft} 
		\item Data-driven error bounds and stopping criteria in \alert{GAIL} \parencite{ChoEtal21a} and \alert{QMCPy} \parencite{QMCPy2020a, ChoEtal22a} 
		\item Uncertainty quantification libraries \alert{Dakota} \parencite{DakotaUsersManual}, \alert{UQTk} \parencite{DebEtal04,UQTk}, and \alert{MUQ} \parencite{MUQ} have some basic level low discrepancy sampling
	    \item QMC framework in Julia since 2019 \parencite{QMCJulia} by Robbe and others
		\item Scrambled Sobol' in \alert{SciPy} \parencite{virtanen2020scipy} and \alert{PyTorch} \parencite{paszke2019pytorch} since several years ago
		\item \alert{TensorFlow} QMC framework \parencite{tfqfQMC2021a} since a year or so ago
	\end{itemize}

In summary, we have stand-alone algorithms, small libraries, and pieces of larger libraries
\end{frame}

@@

\uncover<2->{The United States Department of Energy is investing in studying how to develop great scientific software \parencite{Her19a,ASCR-SSSDU}}
\end{frame}

@@

\begin{frame}{Software Library Architecture}
	\vspace{-7ex}

@@

\vspace{-1ex}
\uncover<2->{May we agree on a \alert{common definition} for these these objects like we have for floating point numbers or basic mathematical functions?}
\end{frame}

\begin{frame}{Choice of Language, Library, or Environment}
	\begin{itemize}
    \item Language/library choices of the users are driven by familiarity and speed

    \item Several favorites aligned with various communities, but no one dominates all

@@

\begin{frame}{Ex.\ Connecting QMCPy with FEniCS/Dolfin \parencite{}}
\vspace{-4ex}
	\begin{itemize}
		\item Want to demonstrate how our QMC software could work with a popular DE solver for DEs with random coefficients
		\item Partial success \only<3->{\alert{Why?}}
	\end{itemize}
 	\vspace{-2ex}
 \begin{minipage}{0.48\textwidth}
 	\only<1-2>{\begin{gather*}
 		-\frac{\dif }{\dif x}\biggl(a(x,\vW) \frac{\dif}{\dif x} u(x,\vW) \biggr) = 200( 2x - 1)\\
 		 0 \leq x \leq 1, \qquad
 		u(0) = u(1) = 0 \\
 		a(x,\vW) = 1 + 0.6 \sum_{k=1}^d \frac{W_k T_{k}(2x-1)}{k^2} \\
 		 \vW \sim \cu[-1,1]^d \\
 		 \Ex[u(0.25,\vW)] = \only<1>{\, ?}\only<2>{ -1.7462}
 	\end{gather*}}
  \only<3->{
\begin{itemize}
\item Even with FEniCS/Dolfin's extensive documentation we had \alert{difficulties}
\begin{itemize}
    \item Learning how to solve a simple ODE, so reverted to an older version
    \item Learning how to change the random instance of coefficient without tiggering the JIT compile
    \item Do not yet know how to express the random coefficient in terms of the covariance kernel
\end{itemize}
\end{itemize}
  }
 \end{minipage}%
\hfill
 \begin{minipage}{0.48\textwidth}
 	\only<1-2>{\centering
 	\includegraphics<1>[width=\textwidth]{ProgramsImages/axwtrue0.6.eps}
 	\includegraphics<2->[width=\textwidth]{ProgramsImages/timing-4.eps}}%
  \only<3->{
\begin{itemize}
\item But
\begin{itemize}
    \item It does work
    \item If we can overcome the challenges, will post a blog and demo at \textcite{QMCBlog}
\end{itemize}
\end{itemize}

>>>>

\usetheme{FJHSlimNoFoot169}
\setlength{\parskip}{2ex}
\setlength{\arraycolsep}{0.5ex}
%\setbeamersize{text margin top=10mm}

@@

\begin{frame}{My Definition of a Great MCQMC Software Ecosystem}

@@

\begin{frame}[allowframebreaks]{Selective History of Publicly Available MC\textbf{QMC} Software}
\vspace{-2ex}
	\begin{itemize}
		\item \textcite{LEc2017a} provides a history of \alert{random number generation}
		\item Notorious \alert{RANDU} from the 1960s \parencite{RANDU}, which fails the spectral test
		\item \alert{VEGAS} \parencite{Lep78a, Lep21a} importance sampling Monte Carlo algorithm favored by physicists
		\item \alert{ACM} low discrepancy point generators \parencite{BraFox88,BraFoxNie92,HonHic00a}
		\item \alert{FinDer} \parencite{PasTra95,FinDer} and \alert{BRODA} \parencite{BRODA20a} targeting quantitative finance
		\item Korobov cubature and scrambled Sobol' generators in \alert{NAG} \parencite{NAG27} for decades		
		\item Scrambled Sobol' and Halton generators in \alert{MATLAB} \parencite{MAT9.13} since 2008, and fixed a few years later
		\item \alert{BUGS} \parencite{BUGSBook, BUGSweb} and \alert{Stan} \parencite{STAN} for Markov Chain Monte Carlo (MCMC)
		\item \alert{LatMRG} \parencite{LEcCou97}, \alert{RNGStreams} \parencite{LEcEtal02},  \alert{SSJ} \parencite{LEc2002a,SSJ}, and \alert{LatNetBuilder} \parencite{LatNet}
		\item Low discrepancy generators \parencite{FriKel02,FriKelweb}, \alert{SamplePack} \parencite{SamplePack}, \textcite{GruWeb},  and \alert{MatBuilder} \parencite{paulin2022}
		\item Sobol' \alert{direction numbers} \parencite{JoeKuo03,JoeKuo08,SobolDirection}
		\item \alert{Fast CBC}, \alert{Magic Point Shop}, \alert{QMC4PDE} and other code since 2004  at \textcite{NuyWeb}
		\item \alert{Multi-level} software \parencite{GilesSoft,GilesQSoft} 
		\item Data-driven error bounds and stopping criteria in \alert{GAIL} \parencite{ChoEtal21a} and \alert{QMCPy} \parencite{QMCPy2020a, ChoEtal22a} 
		\item Uncertainty quantification libraries \alert{Dakota} \parencite{DakotaUsersManual}, \alert{UQTk} \parencite{DebEtal04,UQTk}, and \alert{MUQ} \parencite{MUQ} have some basic level low discrepancy sampling
	    \item QMC framework in Julia since 2019 \parencite{QMCJulia} by Robbe and others
		\item Scrambled Sobol' in \alert{SciPy} \parencite{virtanen2020scipy} and \alert{PyTorch} \parencite{paszke2019pytorch} since several years ago
		\item \alert{TensorFlow} QMC framework \parencite{tfqfQMC2021a} since a year or so ago
	\end{itemize}

We have stand-alone algorithms, small libraries, and pieces of larger libraries
\end{frame}

@@

\uncover<2->{The US Department of Energy is investing in studying how to develop great scientific software \parencite{Her19a,ASCR-SSSDU}}
\end{frame}

@@

\begin{frame}{Software Architecture}
	\vspace{-7ex}

@@

\vspace{-1ex}
\uncover<2->{May we agree on a \alert{common software definition} for these these objects like we have for floating point numbers or basic mathematical functions?}
\end{frame}

\begin{frame}{Choice of Language, Library, or Environment}
	\vspace{-5ex}
	\begin{itemize}
    \item Language/library choices of the users are driven by \alert{familiarity} and \alert{speed}

  	\item  Favorites aligned with various \alert{communities}, but no one dominates all

	\item Difference between \alert{prototyping} environments and ninja written \alert{production} code

@@

\begin{frame}{Ex.\ QMCPy $+$ FEniCS/Dolfin \parencite{LoggEtal_10_2012}}
\vspace{-4ex}
	\begin{itemize}
		\item Want to demonstrate how our QMC software could work with a popular DE solver for DEs with random coefficients
		\item Partial success \only<3->{\alert{Why?}}
	\end{itemize}
 	\vspace{-2ex}
 \begin{minipage}{0.48\textwidth}
 	\only<1-2>{\begin{gather*}
 		-\frac{\dif }{\dif x}\biggl(a(x,\vW) \frac{\dif}{\dif x} u(x,\vW) \biggr) = 200( 2x - 1)\\
 		 0 \leq x \leq 1, \qquad
 		u(0) = u(1) = 0 \\
 		a(x,\vW) = 1 + 0.6 \sum_{k=1}^d \frac{W_k T_{k}(2x-1)}{k^2} \\
 		 \vW \sim \cu[-1,1]^d \\
 		 \Ex[u(0.25,\vW)] = \only<1>{\, ?}\only<2>{ -1.7462}
 	\end{gather*}}
  \only<3->{
\begin{itemize}
\item Even with FEniCS/Dolfin's extensive documentation we had \alert{difficulties}
\begin{itemize}
    \item Learning how to solve a simple ODE, so reverted to an older version
    \item Learning how to change the random instance of coefficient without tiggering the JIT compile
    \item Do not yet know how to express the random coefficient in terms of the covariance kernel
\end{itemize}
\end{itemize}
  }
 \end{minipage}%
\hfill
 \begin{minipage}{0.48\textwidth}
 	\only<1-2>{\centering
 	\includegraphics<1>[width=\textwidth]{ProgramsImages/axwtrue0.6.eps}
 	\includegraphics<2->[width=\textwidth]{ProgramsImages/timing-4.eps}}%
  \only<3->{
\begin{itemize}
\item But
\begin{itemize}
    \item It does work
    \item If we can overcome the challenges, will post a blog and demo at \textcite{QMCBlog}
\end{itemize}
\end{itemize}
